ms_willow_network_supplement

### 2015-04-18 ms_willow_network_supplement.Rmd
made supplement and updated main manuscript

The code below replicates the analyses for the qualitative and quantitative network data.

```{r prepare, echo=FALSE}
## source in required datasets and functions. 
source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
source('~/Documents/miscellaneous_R/model_diagnostic_functions.R')
source('~/Documents/miscellaneous_R/jensen_magnitude_function.R')

## load required libraries
#library(vegan) # for adonis analysis
#library(knitr) # for making tables
library(ggplot2) # for plotting
theme_set(theme_classic()) # customize ggplot for prettier default graphs
library(gridExtra) # for grid.arrange
library(mvabund) # multivariate analysis
library(HH) # for variance inflation factor analysis (vif) in multiple regression

@@

```{r mv abund galls}
gall.mvabund <- mvabund(full.df[ ,c("vLG_abund","rG_abund","aSG_abund","SG_abund")])
meanvar.plot(gall.mvabund ~ full.df$Genotype)
poisson_curve(from = 0.1, to = 10)
quasipoisson_curve(from = 0.1, to = 10, quasi.scalar = 2)
neg.binomial_curve(from = 0.1, to = 10, theta.negbin = 0.7) # negative binomial provides the best fit to the data.
```

Assessed negative binomial assumptions
```{r gall abund negative binomial assumptions}
manyglm.gall <- manyglm(gall.mvabund ~ Genotype,
                        data = full.df,
                        family = "negative.binomial")
plot(manyglm.gall, which = 1:3) # residuals look really good. No heteroscadisticity or non-normality. Note that replotting the residuals gives qualitatively the same picture (it's important to replot them because the residuals involve random number generation, see ?plot.manyglm)
```

Tested whether genotype affected gall community composition
```{r gall abund test}
anova.gall <- anova.manyglm(manyglm.gall, p.uni = "adjusted")
anova.gall # vLG_abund, rG_abund and aSG_abund are driving the community response.
```

Boxplots of gall abundance among the different genotypes.
```{r gall abund across genotypes, echo = FALSE}
vLG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG_abund)) + 
  geom_boxplot() + ylab("Iteomyia salicisverruca density (#/branch)")
rG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG_abund)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides density (#/branch)")
aSG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG_abund)) + 
  geom_boxplot() + ylab("Cecidomyid sp. A density (#/branch)")
SG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG_abund)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbattatus density (#/branch)")

grid.arrange(vLG.abund.plot, rG.abund.plot, aSG.abund.plot, SG.abund.plot)
```

```{r mvabund gall-trait associations}
# selects galls and traits for analysis
galls.traits.df <- full.df %>%
  select(aSG_abund:SG_abund,
         vLG.height.mean, vLG.gall.count,
         Trichome.No., Total_Area, Height, Density,
         C_N_imputed, water_content, specific_leaf_area, sal_tannin.PC1:flavanonOLES.PC1)

@@

```{r stepAIC galls-traits}
## use stepAIC to identify traits that reduce AIC the most and remove them in a stepwise fashion from analysis.
drop1(mvabund.galls.traits)
mod2 <- update(mvabund.galls.traits, .~. - sal_tannin.PC1); drop1(mod2)
mod3 <- update(mod2, .~. - flavonOLES.PC1); drop1(mod3)
mod4 <- update(mod3, .~. -water_content); drop1(mod4)
mod5 <- update(mod4, .~. - cinn.PC2); drop1(mod5)
mod6 <- update(mod5, .~. - cinn.PC1); drop1(mod6)
mod7 <- update(mod6, .~. -SLA_resid); drop1(mod7)
mod8 <- update(mod7, .~. -Height_resid); drop1(mod8)
mod9 <- update(mod8, .~. -Density_resid); drop1(mod9)
mod10 <- update(mod9, .~. -flavonOLES.PC2); drop1(mod10)
mod11 <- update(mod10, .~. -log_trichomes); drop1(mod11)
mod12 <- update(mod11, .~. -C_N_imputed); drop1(mod12)
mod13 <- update(mod12, .~. -log_size); drop1(mod13)
mod14.null <- update(mod13, .~. -flavanonOLES.PC1)

## use AIC to compare models
AIC.models.galls.traits <- data.frame(Model = 1:14, # progressive order in which models were fit
           Formula = c(paste(formula(mvabund.galls.traits)[3]), 
                       paste(formula(mod2)[3]),
                       paste(formula(mod3)[3]), 
                       paste(formula(mod4)[3]),
                       paste(formula(mod5)[3]), 
                       paste(formula(mod6)[3]),
                       paste(formula(mod7)[3]),
                       paste(formula(mod8)[3]), 
                       paste(formula(mod9)[3]),
                       paste(formula(mod10)[3]), 
                       paste(formula(mod11)[3]),
                       paste(formula(mod12)[3]),
                       paste(formula(mod13)[3]),
                       paste(formula(mod14.null)[3])),
           AIC = c(sum(AIC(mvabund.galls.traits)), 
                   sum(AIC(mod2)), 
                   sum(AIC(mod3)), 
                   sum(AIC(mod4)),
                   sum(AIC(mod5)), 
                   sum(AIC(mod6)),
                   sum(AIC(mod7)), 
                   sum(AIC(mod8)), 
                   sum(AIC(mod9)),
                   sum(AIC(mod10)), 
                   sum(AIC(mod11)),
                   sum(AIC(mod12)), 
                   sum(AIC(mod13)), 
                   sum(AIC(mod14.null))))
arrange(AIC.models.galls.traits, AIC)

## Likelihood ratio tests suggest that all of the models with lower AIC than the null model provide a better fit to the data than the null.
anova.manyglm(mod14.null, mod13)
anova.manyglm(mod14.null, mod12)
anova.manyglm(mod14.null, mod11)
anova.manyglm(mod14.null, mod10)

# identify gall-trait associations. C_N_imputed has a marginally significant effect on Iteomyia and R. salicisbrassicoides abundance. FlavononOLES.PC1 has a signficiant effect on Cecidomyiid sp. A abundance, and plant size as a significant effect on R. salicisbrassicoides and a marginal effect on R. salicisbattatus abundance.
uni.galls.traits <-anova.manyglm(mod11, p.uni = "unadjusted")

@@

```{r visreg univariate gall-trait associations}
library(visreg)
# include model with all of the traits, but only examine one of the variables
vLG.trait <- MASS::glm.nb(vLG_abund ~ C_N_imputed + log_size + flavanonOLES.PC1, gall.density.df)
summary(vLG.trait) 
visreg(vLG.trait, xvar = "C_N_imputed", scale = "response")

@@

```{r}
## gall size
gall.size.df <- galls.traits.df %>%
  mutate(log_size = log(Total_Area),
         log_trichomes = log(Trichome.No.+1)) %>%
  select(vLG.height.mean, vLG.gall.count, Height:log_trichomes) %>%
  na.omit()
gall.size.df$Height_resid = resid(lm(Height ~ log_size, gall.size.df))
gall.size.df$Density_resid = resid(lm(Density ~ log_size, gall.size.df))
gall.size.df$SLA_resid = resid(lm(specific_leaf_area ~ water_content, gall.size.df))
gall.size.df <- select(gall.size.df, vLG.height.mean, vLG.gall.count, 
                       C_N_imputed, water_content,
                       sal_tannin.PC1:SLA_resid)

@@

# all of these models give the same picture. Sal_tannin.PC1 and flavonOLES.PC1 are the best predictors.
summary(size11)
summary(size10)
summary(size12)
summary(size9)

@@

```{r gall abund predict}
gall.predict <- predict(manyglm.gall, 
                        newdata = data.frame(Genotype = levels(full.df$Genotype)), 
                        type = "response")
gall.predict.df <- cbind.data.frame(gall.predict, Genotype = levels(full.df$Genotype))
gall.predict.melt <- melt(gall.predict.df)

ggplot(gall.predict.melt, aes(x = Genotype, y = value, fill = variable)) + geom_bar(stat = "identity")

colMeans(gall.predict)
library(vegan)
diversity(colMeans(gall.predict))

gall.predict.X.R <- predict(manyglm.gall, 
                            newdata = data.frame(Genotype = c("X","R")),
                            type = "response")
colMeans(gall.predict.X.R)
diversity(colMeans(gall.predict.X.R))

sample(levels(full.df$Genotype), 5)

names(manyglm.gall$data)
predict.allcombos <- function(manyglm.object, factor.index, predict.type = "response"){
  require(plyr)
  require(dplyr)
  factor.levels <- levels(manyglm.object$data[ ,factor.index])

  predict.list <- list()
  for(i in 2:length(factor.levels)){
    tmp.data <- data.frame(sample(factor.levels, i))
    colnames(tmp.data) <- names(manyglm.object$data)[factor.index]
    predict.tmp <- predict(manyglm.object, newdata = tmp.data, type = predict.type)
    predict.list[[i]] <- colMeans(predict.tmp)
  }
  predict.df <- ldply(predict.list) %>%
    mutate(Number.of.Genotypes = 2:length(factor.levels))
}

N <- 50
gall.div.list <- list()
for(i in 1:N){
  gall.div.list[[i]] <- round(predict.allcombos(manyglm.gall, factor.index = 2), 1) # rounding to avoid pseudoaccuracy
} # takes a couple minutes with 50 reps

gall.div.df <- ldply(gall.div.list) %>% tbl_df()
gall.div.df$diversity <- diversity(select(gall.div.df, vLG_abund:SG_abund))

boxplot(diversity ~ factor(Number.of.Genotypes), gall.div.df, ylab = "Gall diversity", xlab = "Number of Genotypes")

N <- 50
link.div.list <- list()
for(i in 1:N){
  link.div.list[[i]] <- round(predict.allcombos(manyglm.full, factor.index = 2), 1)
} # takes a couple minutes with 50 reps

link.div.df <- ldply(link.div.list) %>% tbl_df()
link.div.df$diversity <- diversity(select(link.div.df, aSG_Tory:vLG_Tory))
link.div.df$repID <- 1:dim(link.div.df)[1]

boxplot(diversity ~ factor(Number.of.Genotypes), link.div.df, ylab = "Interaction Diversity", xlab = "Number of Genotypes")

link.melt <- melt(select(link.div.df, -diversity), id.vars = c("repID", "Number.of.Genotypes"))

link.split <- colsplit(link.melt$variable, "_", names = c("gall","parasitoid"))
link.melt.df <- cbind(link.melt, link.split)

web.list <- cast(link.melt.df, gall ~ parasitoid | repID + Number.of.Genotypes, mean)

library(bipartite)
web.list.measures <- c()
genotype.number <- c()
for(i in 1:dim(web.list)){
  genotype.number[i] <- names(web.list[[i]][1])
  tmp.web <- web.list[[i]][[1]]
  rownames(tmp.web) <- tmp.web$gall
  tmp.web <- as.matrix.data.frame(tmp.web[ ,-1])
  web.list.measures[i] <- networklevel(tmp.web, index = "linkage density", weighted = TRUE)
}
web.measures.df <- cbind.data.frame(Number.of.Genotypes = as.numeric(genotype.number),
                                    Weighted.Linkage.Density = web.list.measures)

boxplot(Weighted.Linkage.Density ~ factor(Number.of.Genotypes), web.measures.df, ylab = "Weighted Linkage Density", xlab = "Number of Genotypes")

web.measures.summary <- web.measures.df %>%
  group_by(Number.of.Genotypes) %>%
  summarise(median.link.density = median(Weighted.Linkage.Density))
with(web.measures.summary, max(median.link.density)/min(median.link.density))

library(mgcv)
web.sum.lm <- lm(log(median.link.density) ~ log(Number.of.Genotypes), data = web.measures.summary)
summary(web.sum.lm) # Need to understand interpretation here.
visreg(web.sum.lm) # not quite there
```

We were unable to use the mvabund framework for this analysis, so we conducted separate linear models to test for these effects.
```{r gall size}
plot(vLG.height.mean ~ Genotype, full.df)
vLG.size.lm <- lm(log(vLG.height.mean) ~ Genotype, filter(full.df, Genotype != "J"))
#plot(vLG.size.lm)
summary(vLG.size.lm)

vLG.size.lmer <- lmer(log(vLG.height.mean) ~ (1|Genotype), full.df)
heritability(vLG.size.lmer) # significant. Note though, that I originally valuate gall size using nested random effect models to account for individual measurements I had for each gall.

plot(rG.height.mean ~ Genotype, full.df)
rG.size.lm <- lm(log(rG.height.mean) ~ Genotype, full.df)
#plot(rG.size.lm)
summary(rG.size.lm)

plot(aSG.height.mean ~ Genotype, full.df) # too few data points to reliably evaluate
aSG.size.lm <- lm(log(aSG.height.mean) ~ Genotype, full.df)
#plot(aSG.size.lm)
summary(aSG.size.lm)

plot(SG.height.mean ~ Genotype, full.df) # too few data points to reliably evaluate
SG.size.lm <- lm(log(SG.height.mean) ~ Genotype, full.df)
plot(SG.size.lm)
summary(SG.size.lm)
```

```{r gall size variation among willow genotypes}
vLG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG.height.mean)) + 
  geom_boxplot() + ylab("Iteomyia salicisverruca diameter (mm)")
rG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG.height.mean)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides diameter (mm)")
aSG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG.height.mean)) + 
  geom_boxplot() + ylab("Cecidomyid sp. A diameter (mm)")
SG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG.height.mean)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbattatus diameter (mm)")

Here, we evaluate the assumptions of mvabund and see which error distribution is appropriate. Specifically, we first look at a plot of the mean-variance relationship of our response variables. It is easy to see that the negative binomial model (black line) provides the best fit to this data, suggesting that we should specify this as the error distribution in our model
```{r mv abund interactions explore}
full.mvabund <- mvabund(full.df[ ,interaxns_noPont]) # create mvabund object for easy analysis
meanvar.plot(full.mvabund ~ full.df$Genotype) # easy to see that the data do not have a poisson distribution, therefore, a negative binomial may be a good fit.
poisson_curve(from = 0.1, to = 3.0, color = "blue")
quasipoisson_curve(from = 0.1, to = 3.0, quasi.scalar = 2, color = "red")
neg.binomial_curve(from = 0.1, to = 3.0, theta.negbin = 0.7, color = "black") # negative binomial appears to provide the best fit to this relationship
```

```{r genotype level}
geno.level.interactions.df <- full.df %>%
  select(Genotype, aSG_Tory:rG_Platy, rG_Tory, SG_Platy, vLG_Eulo:vLG_Tory, aSG_abund:vLG_abund, vLG.height.mean) %>%
  group_by(Genotype) %>%
  summarise_each(funs(mean.na.rm = mean(., na.rm = TRUE))) %>%
  #na.omit() %>%
  mutate(log.vLG_abund = log(vLG_abund),
         log.vLG.height.mean = log(vLG.height.mean),
         log.1.rG_abund = log(rG_abund+1),
         log.1.aSG_abund = log(aSG_abund+1))

library(psych)
corr.test(geno.level.interactions.df[ ,c("vLG_abund","rG_abund","aSG_abund","vLG.height.mean")])

car::scatterplotMatrix(select(geno.level.interactions.df, vLG_abund, rG_abund, aSG_abund, vLG.height.mean))
corr.test(geno.level.interactions.df[ ,c("log.vLG_abund","log.1.rG_abund","log.1.aSG_abund","log.vLG.height.mean")])

# correlations are rather weak
corr.test(select(geno.level.interactions.df, aSG_Tory:vLG_Tory))
corr.test(select(geno.level.interactions.df, vLG_Platy, vLG_Mesopol, vLG_Tory, rG_Tory))
car::scatterplotMatrix(select(geno.level.interactions.df, vLG_Platy, vLG_Mesopol, vLG_Tory, rG_Tory))

plot(rG_Eulo ~ vLG_Mymarid, geno.level.interactions.df)

plot(log.1.aSG_abund ~ log.vLG_abund, geno.level.interactions.df)

geno.level.mvabund <- mvabund(geno.level.interactions.df[ ,interaxns_noPont])

geno.level.manyglm <- manylm(geno.level.mvabund ~ vLG_abund*vLG.height.mean + rG_abund + aSG_abund, data = geno.level.interactions.df)

plot(geno.level.manyglm, which = 1:3)
anova.manylm(geno.level.manyglm, p.uni = "unadjusted")

hist(geno.level.interactions.df$vLG_Platy)
plot(vLG_Mesopol ~ vLG.height.mean, geno.level.interactions.df)
vLG_Platy.lm <- lm(vLG_Mesopol ~ vLG_abund+vLG.height.mean, geno.level.interactions.df)
summary(vLG_Platy.lm)
plot(vLG_Platy.lm)
AIC(vLG_Platy.lm)
library(visreg)
visreg(vLG_Platy.lm, scale = "response")
```

@@

```{r}
link.predict <- predict(manyglm.full, 
                        newdata = data.frame(Genotype = levels(full.df$Genotype)), 
                        type = "response")
link.predict.df <- cbind.data.frame(link.predict, Genotype = levels(full.df$Genotype))
library(vegan)
mean(vegdist(select(link.predict.df, -Genotype), method = "horn"))
sd(vegdist(select(link.predict.df, -Genotype), method = "horn"))
hist(vegdist(select(link.predict.df, -Genotype), method = "horn"))

link.predict.melt <- melt(link.predict.df)

ggplot(link.predict.melt, aes(x = Genotype, y = value, fill = variable)) + geom_bar(stat = "identity")
```

@@

From the table, it is clear to say that genotype has a strong effect on the composition of links in the network. Moreover, differences in community composition are driven primarily by 3 interactions: vLG_Platy, vLG_Tory, and vLG_Mesopol.
```{r}
anova.full <- anova.manyglm(manyglm.full, p.uni = "unadjusted") # Takes about 1 min and 30 sec to run.
anova.full # vLG_Platy, vLG_Tory, and vLG_Mesopol are driving the community response. rG_Tory is marginally significant and vLG_Eulo is close to marginal as well.
```

@@

We then examined the which processes underlied variation in the gall-parasitoid interactions.
```{r}
anova.net <- anova.manyglm(net.mvabund.2, p.uni = "unadjusted") # takes about 2.5 min to run.
anova.net # not that since the P-values are determined by a resampling procedure, they may differ slightly between runs. Therefore, we retain all P-values < 0.10 (during at least one run) for the coefficient summary below.

>>>>

```{r prepare, include = FALSE, echo=FALSE}
## source in required datasets and functions. 
source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
source('~/Documents/miscellaneous_R/model_diagnostic_functions.R')
#source('~/Documents/miscellaneous_R/jensen_magnitude_function.R')

## load required libraries
library(vegan) # for adonis analysis
library(knitr) # for making tables
library(ggplot2) # for plotting
theme_set(theme_bw()) # customize ggplot for prettier default graphs
library(gridExtra) # for grid.arrange
library(mvabund) # multivariate analysis
library(HH) # for variance inflation factor analysis (vif) in multiple regression
library(visreg) # visualize regression output

@@

Testing for whether gall size varies among willow genotypes. We were unable to use the mvabund framework for this analysis, so we conducted separate linear models to test for these effects.

```{r gall size plots, echo=FALSE}
vLG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG.height.mean)) + 
  geom_boxplot() + ylab("Iteomyia salicisverruca diameter (mm)")
rG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG.height.mean)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides diameter (mm)")
aSG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG.height.mean)) + 
  geom_boxplot() + ylab("Cecidomyid sp. A diameter (mm)")
SG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG.height.mean)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbattatus diameter (mm)")

There were too few data points to reliably test for differences in mid-Stem gall diameter (Rabdophaga salicisbattatus), but we used weighted linear models to test for differences in gall size among willow genotypes. We weighted models by the number of galls collected from each plant replicate, because a higher number would reflect a more accurate measure of mean gall size for that particular willow.

```{r gall size models}
vLG.size.lm <- lm(log(vLG.height.mean) ~ Genotype, data = full.df, weights = vLG.gall.count)
rG.size.lm <- lm(log(rG.height.mean) ~ Genotype, data = full.df, weights = rG.gall.count)
aSG.size.lm <- lm(log(aSG.height.mean) ~ Genotype, data = full.df, weights = aSG.gall.count)
```

```{r gall size results summary, echo=FALSE}
anova(vLG.size.lm)
anova(rG.size.lm)
anova(aSG.size.lm)
```

We used multivariate GLMs to examine whether gall abundances varied among willow genotypes.

Boxplots of gall abundance among the different genotypes.
```{r gall abund across genotypes, echo = FALSE}
gall.plot.df <- as.data.frame(full.df) %>% # melt doesn't like 'tbl' class
  select(Genotype, 
         Iteomyia_salicisverruca = vLG_abund,
         Cecidomyiid_sp.A = aSG_abund,
         Rabdophaga_salicisbrassicoides = rG_abund, 
         Rabdophaga_salicisbattatus = SG_abund) %>%
  melt()

ggplot(gall.plot.df, aes(x = Genotype, y = value, color = variable)) + 
  geom_boxplot() + facet_wrap( ~ variable, nrow = 2) + theme_bw() + ylab("Gall density (no. per branch)")
```

Assessing which error distribution is best to model the data. Negative binomial appear to provide the best fit.
```{r gall abundance negative binomial assumption, echo=FALSE}
gall.mvabund <- mvabund(full.df[ ,c("vLG_abund","rG_abund","aSG_abund","SG_abund")])
meanvar.plot(gall.mvabund ~ full.df$Genotype)
poisson_curve(from = 0.1, to = 10)
quasipoisson_curve(from = 0.1, to = 10, quasi.scalar = 2)
neg.binomial_curve(from = 0.1, to = 10, theta.negbin = 0.7) # negative binomial

## specify the multivariate GLM
manyglm.gall <- manyglm(gall.mvabund ~ Genotype,
                        data = full.df,
                        family = "negative.binomial")

## examine residual plots of the model.
plot(manyglm.gall, which = 1:3) # residuals look really good. No heteroscadisticity or non-normality. Note that replotting the residuals gives qualitatively the same picture (it's important to replot them because the residuals involve random number generation, see ?plot.manyglm)
```

Tested whether genotype affected gall community composition. It appears that leaf and bud galls are the major determinants of variation in gall abundance, with apical-Stem galls contributing to a lesser extent. Remember that p-values may slightly differ from those reported in the manuscript since this uses a permutation.
```{r gall abund test, eval=FALSE}
anova.gall <- anova.manyglm(manyglm.gall, p.uni = "unadjusted")
anova.gall 
```

```{r test gall linear models}
gall.lins <- manylm(log(gall.mvabund+1) ~ Genotype, data = full.df)
summary(gall.lins)
anova(gall.lins, p.uni = "unadjusted") # same qualitative results as GLM
```

Test for quantitative differences  in gall community composition.
```{r vegan galls, echo=FALSE}
gall.comm <- full.df %>% select(Genotype, vLG_abund, rG_abund, aSG_abund, SG_abund)
gall.comm.sub <- filter(gall.comm, rowSums(gall.comm[ ,-1]) > 0, 
                        Genotype != "J", Genotype != "N", Genotype != "U")
table(gall.comm.sub$Genotype) # restricted to Genotypes with greater than 2 replicates for composition analyses.

bray.gall.comm.sub <- vegdist(gall.comm.sub[ ,-1], method = "bray")

adonis(bray.gall.comm.sub ~ Genotype, data = gall.comm.sub)
anova(betadisper(bray.gall.comm.sub, group = gall.comm.sub$Genotype)) # no differences in dispersion among willow genotypes.
summary(meandist(bray.gall.comm.sub, grouping = gall.comm.sub$Genotype))
```

Determine which willow traits are resulting in variation in gall abundances.
```{r mvabund gall-trait associations, echo=FALSE}
# selects galls and traits for analysis
galls.traits.df <- full.df %>%
  select(aSG_abund:SG_abund,
         vLG.height.mean, vLG.gall.count,
         Trichome.No., Total_Area, Height, Density,
         C_N_imputed, water_content, specific_leaf_area, sal_tannin.PC1:flavanonOLES.PC1)

@@

Starting with the most complex model, we progressively dropped predictors from the model that resulted in the lowest AIC for the model. We continued this in stepwise fashion until we arrived at the null model (no predictors). Using AIC and likelihood ratio tests, we found that four of these models provided a better fit to the data than the null model. For clarity, we only interpret model 11 as this one had the second to lowest AIC value, yet likelihood ratio tests suggested that it provided a significantly better fit than the alternative models. 
```{r stepAIC galls-traits, echo=FALSE, results='asis'}
## drop variables that reduce AIC the most and remove them in a stepwise fashion from analysis.
drop1(mvabund.galls.traits)
mod2 <- update(mvabund.galls.traits, .~. - sal_tannin.PC1); drop1(mod2)
mod3 <- update(mod2, .~. - flavonOLES.PC1); drop1(mod3)
mod4 <- update(mod3, .~. -water_content); drop1(mod4)
mod5 <- update(mod4, .~. - cinn.PC2); drop1(mod5)
mod6 <- update(mod5, .~. - cinn.PC1); drop1(mod6)
mod7 <- update(mod6, .~. -SLA_resid); drop1(mod7)
mod8 <- update(mod7, .~. -Height_resid); drop1(mod8)
mod9 <- update(mod8, .~. -Density_resid); drop1(mod9)
mod10 <- update(mod9, .~. -flavonOLES.PC2); drop1(mod10)
mod11 <- update(mod10, .~. -log_trichomes); drop1(mod11)
mod12 <- update(mod11, .~. -C_N_imputed); drop1(mod12)
mod13 <- update(mod12, .~. -log_size); drop1(mod13)
mod14.null <- update(mod13, .~. -flavanonOLES.PC1)

## use AIC to compare models
AIC.models.galls.traits <- data.frame(Model = 1:14, # progressive order in which models were fit
           Formula = c(paste(formula(mvabund.galls.traits)[3]), 
                       paste(formula(mod2)[3]),
                       paste(formula(mod3)[3]), 
                       paste(formula(mod4)[3]),
                       paste(formula(mod5)[3]), 
                       paste(formula(mod6)[3]),
                       paste(formula(mod7)[3]),
                       paste(formula(mod8)[3]), 
                       paste(formula(mod9)[3]),
                       paste(formula(mod10)[3]), 
                       paste(formula(mod11)[3]),
                       paste(formula(mod12)[3]),
                       paste(formula(mod13)[3]),
                       paste(formula(mod14.null)[3])),
           AIC = c(sum(AIC(mvabund.galls.traits)), 
                   sum(AIC(mod2)), 
                   sum(AIC(mod3)), 
                   sum(AIC(mod4)),
                   sum(AIC(mod5)), 
                   sum(AIC(mod6)),
                   sum(AIC(mod7)), 
                   sum(AIC(mod8)), 
                   sum(AIC(mod9)),
                   sum(AIC(mod10)), 
                   sum(AIC(mod11)),
                   sum(AIC(mod12)), 
                   sum(AIC(mod13)), 
                   sum(AIC(mod14.null))))
kable(arrange(AIC.models.galls.traits, AIC)[1:5,]) # display all of the models that had lower AICs than the null model (including the null model).

## Likelihood ratio tests suggest that all of the models with a lower AIC than the null model provide a better fit to the data than the null.
anova.manyglm(mod14.null, mod13)
anova.manyglm(mod14.null, mod12)
anova.manyglm(mod14.null, mod11)
anova.manyglm(mod14.null, mod10)

## Likelihood ratio tests suggest that model 11 provides a better fit to the data than the two other closest models
anova.manyglm(mod13, mod11)
anova.manyglm(mod12, mod11)
anova.manyglm(mod10, mod11)
anova.manyglm(mod14.null, mod13, mod12, mod11, mod10)
```

```{r identify gall-trait associations, eval=FALSE}
# identify gall-trait associations. C_N_imputed has a marginally significant effect on Iteomyia and R. salicisbrassicoides abundance. FlavononOLES.PC1 has a signficiant effect on Cecidomyiid sp. A abundance, and plant size as a significant effect on R. salicisbrassicoides and a marginal effect on R. salicisbattatus abundance.
uni.galls.traits <-anova.manyglm(mod11, p.uni = "unadjusted")

@@

```{r gall trait lms}
gall.trait.lms <- manylm(galls_galls.traits ~ C_N_imputed + flavanonOLES.PC1 + log_size, 
                                data = traits_galls.traits)
summary(gall.trait.lms)
anova(gall.trait.lms, p.uni = "unadjusted") # qualitatively the same results as GLM.

best.r.sq(galls_galls.traits ~ traits_galls.traits$C_N_imputed + traits_galls.traits$water_content + traits_galls.traits$sal_tannin.PC1 + traits_galls.traits$cinn.PC1 + traits_galls.traits$cinn.PC2 + traits_galls.traits$flavonOLES.PC1 + traits_galls.traits$flavonOLES.PC2 + traits_galls.traits$flavanonOLES.PC1 + traits_galls.traits$log_size + traits_galls.traits$log_trichomes + traits_galls.traits$Height_resid + traits_galls.traits$Density_resid + traits_galls.traits$SLA_resid) # results are pretty close to GLM, retains flavanonOLES.PC1, C:N, but has cinn.PC1 and log.trichomes before log.size. Note that these variables were included in the GLM AIC, so it isn't too surprising. Not a big concern overall though.
```

```{r visreg univariate gall-trait associations, eval=FALSE}

# include model with all of the traits, but only examine one of the variables
vLG.trait <- MASS::glm.nb(vLG_abund ~ C_N_imputed + log_size + flavanonOLES.PC1, gall.density.df)
summary(vLG.trait) 
visreg(vLG.trait, xvar = "C_N_imputed", scale = "response")

@@

```{r gall size trait associations, echo=FALSE}
## gall size
gall.size.df <- galls.traits.df %>%
  mutate(log_size = log(Total_Area),
         log_trichomes = log(Trichome.No.+1)) %>%
  select(vLG.height.mean, vLG.gall.count, Height:log_trichomes) %>%
  na.omit()
gall.size.df$Height_resid = resid(lm(Height ~ log_size, gall.size.df))
gall.size.df$Density_resid = resid(lm(Density ~ log_size, gall.size.df))
gall.size.df$SLA_resid = resid(lm(specific_leaf_area ~ water_content, gall.size.df))
gall.size.df <- select(gall.size.df, vLG.height.mean, vLG.gall.count, 
                       C_N_imputed, water_content,
                       sal_tannin.PC1:SLA_resid)

@@

# all of these models give the same picture. Sal_tannin.PC1 and flavonOLES.PC1 are the best predictors.
summary(size11)
summary(size10)
summary(size12) # final model
summary(size9)

summary(lm(vLG.height.mean ~ luteolin.der2__A320nm, full.df, weights = full.df$vLG.gall.count))

anova(size14.null, size13, size12, size11, size10, size9, size8, size7)

@@

Here, we evaluate the assumptions of mvabund and see which error distribution is appropriate. Specifically, we first look at a plot of the mean-variance relationship of our response variables. It is easy to see that the negative binomial model (black line) provides the best fit to this data, suggesting that we should specify this as the error distribution in our model
```{r mv abund interactions explore, echo=FALSE}
full.mvabund <- mvabund(full.df[ ,interaxns_noPont]) # create mvabund object for easy analysis
full.meanvar <- meanvar.plot(full.mvabund ~ full.df$Genotype, table = TRUE) # easy to see that the data do not have a poisson distribution, therefore, a negative binomial may be a good fit.
poisson_curve(from = 0.1, to = 3.0, color = "blue")
quasipoisson_curve(from = 0.1, to = 3.0, quasi.scalar = 2, color = "red")
neg.binomial_curve(from = 0.1, to = 3.0, theta.negbin = 0.7, color = "black") # negative binomial appears to provide the best fit to this relationship

#no.relationships <- which(names(colSums(full.mvabund)) %in% c("rG_Lestodip", "rG_Mesopol", "SG_Platy", "vLG_Eulo"))
#sum(colSums(full.mvabund)[no.relationships])/sum(colSums(full.mvabund)) # less than 13% of the links in the network had relationships to gall abundance and size (see results below)
```

@@

@@

From the table, it is clear to say that genotype has a strong effect on the composition of links in the network. Moreover, differences in community composition are driven primarily by 3 interactions: vLG_Platy, vLG_Tory, and vLG_Mesopol.
```{r, eval=FALSE}
anova.full <- anova.manyglm(manyglm.full, p.uni = "unadjusted") # Takes about 1 min and 30 sec to run.
anova.full # vLG_Platy, vLG_Tory, and vLG_Mesopol are driving the community response. rG_Tory is marginally significant and vLG_Eulo is close to marginal as well. aSG_Tory is significant as well.
```

```{r test linear models link abundance}
link.lins <- manylm(log(full.mvabund+1) ~ Genotype, data = full.df)
summary(link.lins)
anova(link.lins, p.uni = "unadjusted") # qualitatively same results as GLM
```

```{r composition analysis of gall-parasitoid links, echo=FALSE}
link.comm <- cbind.data.frame(Genotype = full.df$Genotype, full.df[ ,interaxns_noPont])

## filtered the data so that all willows have at least one trophic interaction (necessary for quantifying dissimilarity) and that the Genotypes included have 3 or more replicates
link.comm.sub <- filter(link.comm, rowSums(link.comm[ ,-1]) > 0, # must have at least one trophic interaction
                        Genotype %in% c("*","B", "D", "I", "K", "L", 
                                        "Q", "S", "V", "W", "X", "Y", "Z")) 
table(link.comm.sub$Genotype)

bray.link.comm.sub <- vegdist(link.comm.sub[ ,-1], method = "bray")

adonis(bray.link.comm.sub ~ Genotype, data = link.comm.sub)
anova(betadisper(bray.link.comm.sub, group = link.comm.sub$Genotype)) # no differences in dispersion among willow genotypes.
summary(meandist(bray.link.comm.sub, grouping = link.comm.sub$Genotype))
```

@@

We then examined the which processes underlied variation in the gall-parasitoid interactions.
```{r, eval=FALSE}
anova.net <- anova.manyglm(net.mvabund.2, p.uni = "unadjusted") # takes about 2.5 min to run.
anova.net # not that since the P-values are determined by a resampling procedure, they may differ slightly between runs. Therefore, we retain all P-values < 0.10 (during at least one run) for the coefficient summary below.

t(coef(net.mvabund.2))[,-1][c(5,4,6,2,3,1,7),c(2,1,3,4)]

anova.net.null <- anova.manyglm(net.mvabund.2, net.mvabund.6.null) # takes about 30 sec to run
anova.net.null
